---
title: "Creating Visualizations with ggplot"
author: "Benjamin Cassidy"
date: "1/27/2020"
output: html_document
---

## Plots with ggplot2

We are going to make plots in R using the ggplot2 package from the tidyverse.


## Methods:
We will install these **packages**

 - **tidyverse** : suite of packages for wrangling and viz with ggplot
 - **readxl** : to read in Excels into R
 - **here** : to help with filepath syntax

```{r setup, message=FALSE}
library(tidyverse) # install.packages("tidyverse")
library(readxl) # install.packages("readxl")
library(here) # install.packages("here")


```

## Data:
Here we will read in the data from the **California National Parks** and **California Channel Islands** data sets
```{r read-data}

ca_np <- read_csv(here("data/ca_np.csv"))
#alternative code -- omit 'here' but then will be filepath specific
#alternative code 2 -- nere("data","ca_np.csv")

ci_np <- read_xlsx(here("data/ci_np.xlsx"))
#alternative code -- call read_excel instead


```

## Visualization: 

Using ggplot2 to make visualizaions from the data -- 

Makes these layer by layer to create a code layer and map them for repeatability.

Here are the steps: 
 - Tell R you're using ggplot
 - A: What data frame are you getting the variables from?
 - A: What variable(s) do you want to plot?
 - A: What type of graph do you want to make using (geom_)?
 
### Example of ggplot line plot for ci_np vistors X year

```{r}
# X == year
# Y == vistors

ggplot(data = ci_np, aes(x = ci_np$year, y = ci_np$visitors)) + geom_line()



```

```{r}
# making base layer for ci_np viz

gg_base <- ggplot(data = ci_np, aes(x = ci_np$year, y = ci_np$visitors))

gg_base + geom_point()

```


### Customization of visualizations

Here we will learn about how to customize elements of the visualizations

We can update:
 - Line or point colors: color
 - Change line type: linetype
 - Change fill: fill 


```{r}
gg_base + 
  geom_point(
    color = "dodgerblue",
    pch = 13, 
    size = 3
    
  )
# updating based on constants gives the output below

```

### Update the graph aesthetics based on variables

Here we will create changes based on the continuous variable of **vistors**


```{r}

gg_base + 
  geom_point(
    aes(size = ci_np$visitors, 
        color = ci_np$year), 
    pch = 8
        
  )

```

### Usings themes in ggplot via gg-themes for overall updates

Use the following recommended themes as examples: 
 - theme_light
 - theme_minimal
 - theme_bw

```{r}

gg_base +
  geom_area() +
  theme_linedraw()

```


### Visualization in series - Moving on to the **ca_np** dataset 

Will give ggplot the whole set and let it figure out the breakpoints and the rest

```{r}

# a bad graph that looks cool -- not respective of the groupings
gg_base2 <- ggplot(data = ca_np, aes(x=ca_np$year, y=ca_np$visitors))

gg_base2 + geom_line()

# adding in the group by 

gg_base3 <- ggplot(data = ca_np, aes(x=ca_np$year, y=ca_np$visitors, group= park_name))

gg_base3 + 
  geom_line( aes(color = park_name) ) + 
              theme_minimal()
  



```

### Faceting via ggplot to break out graphs

```{r}
my_np_graph <- gg_base2 + 
  geom_point(aes(color =park_name), show.legend = FALSE) +
  facet_wrap(~park_name, scales = "free") + 
  theme_linedraw()

#scales default to fixed (all same), so can jailbreak it with 'free'
#legend default to show unless show.lengend= FALSE

my_np_graph


```

### Export my graph as a .png file

```{r}

ggsave(plot= my_np_graph, here("figures/np_graph1.png"))



```



## Pivot Tables - pots-ability with Lawbsta data
















