---
title: "Creating Visualizations with ggplot"
author: "Benjamin Cassidy"
date: "1/27/2020"
output: html_document
---

## Plots with ggplot2

We are going to make plots in R using the ggplot2 package from the tidyverse.


## Methods:
We will install these **packages**

 - **tidyverse** : suite of packages for wrangling and viz with ggplot
 - **readxl** : to read in Excels into R
 - **here** : to help with filepath syntax

```{r setup, message=FALSE}
library(tidyverse) # install.packages("tidyverse")
library(readxl) # install.packages("readxl")
library(here) # install.packages("here")


```

## Data:
Here we will read in the data from the **California National Parks** and **California Channel Islands** data sets
```{r read-data}

ca_np <- read_csv(here("data/ca_np.csv"))
#alternative code -- omit 'here' but then will be filepath specific
#alternative code 2 -- nere("data","ca_np.csv")

ci_np <- read_xlsx(here("data/ci_np.xlsx"))
#alternative code -- call read_excel instead


```

## Visualization: 

Using ggplot2 to make visualizaions from the data -- 

Makes these layer by layer to create a code layer and map them for repeatability.

Here are the steps: 
 - Tell R you're using ggplot
 - A: What data frame are you getting the variables from?
 - A: What variable(s) do you want to plot?
 - A: What type of graph do you want to make using (geom_)?
 
### Example of ggplot line plot for ci_np vistors X year

```{r}
# X == year
# Y == vistors

ggplot(data = ci_np, aes(x = ci_np$year, y = ci_np$visitors)) + geom_line()



```


