---
title: "Tidying Inverts"
author: "Benjamin Cassidy"
date: "1/28/2020"
output: html_document
---


## Set-up Documentation - Packages
```{r package_installs, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)


library(tidyverse) #install.packages("tidyverse")
library(janitor) #install.packages("janitor")
library(readxl) #install.packages("readxl")
library(here) #install.packages("here")
#library(skimr) #install.packages("skimr")
#library(kableExtra) #install.packages("kableExtra")


```


## Data

Inverts Data will be read in here

```{r data_load}

inverts <- read_xlsx(here("data","inverts.xlsx"))

head(inverts)

```




## Tidying data with `pivot_longer()`
```{r pivot_longer_table}

inverts_long <- pivot_longer(data = inverts,
             cols = c('2016','2017','2018'), #alternative code -- cols = '2016':'2018' to call in series
             names_to = 'year',
             values_to = 'obs_count')

inverts_long

```

## Tidying data with `pivot_wider()`
```{r pivot_wider_table}

inverts_wide <- inverts_long %>% 
  pivot_wider(names_from = common_name, 
                     values_from = obs_count)

inverts_wide

```


## Tidying data with `janitor::clean_names()`

Remove spaces from the column names. 

Defaults to lowercase snakecase columns


```{r janitor_clean_names}

inverts_wide <- inverts_wide %>% janitor::clean_names()

inverts_wide



```



## Tidying data with `tidyr::unite()`

Combines information from columns in a table

```{r unite_site_year}

#create a single column called site_year

inverts_unite <- inverts_long %>% unite(col = "site_year",
                                        c(site,year),
                                        sep = "_")


inverts_unite

```


### EXAMPLE - Creating a month-year column with a '/' separator

```{r mon_year_ex}

inverts_moyr <- inverts_long %>% unite(col = "mo_yr",
                                        c(year,month),
                                        sep = "/") 

inverts_moyr


#with fun separator

inverts_moyr2 <- inverts_long %>% unite(col = "mo_yr",
                                        c(year,month),
                                        sep = "..^o__o^..") 
inverts_moyr2



```



## Pulling part data with `tidyr::separate()`

Start with a combined column to create some additional column

```{r pull_apart_data}

inverts_sep <- inverts_unite %>% tidyr::separate(col = "site_year",
                                                 into = c('my_site','my_year'),
                                                 sep = '_')



inverts_sep

```


## Taking notes on a criminal conspiracy with `stringR` Bell


Use `diplyr::mutate` and `stringR`


### Column mutation
Using `trees` as an example dataset

```{r trees_high_five}
# create new column

tree_five_on_it <- trees %>% mutate(high_five = Height + 5)

tree_five_on_it

# mutate does overwrite existing columns

tree_five_effd <- tree_five_on_it %>% mutate(Height = Height + 10)

tree_five_effd


```

### StringR `str_replace()` 

Replacing column values 

```{r ca_fishes}

ca_abbr <- inverts %>% mutate(common_name = 
                                str_replace(common_name, 
                                            pattern = "california",
                                            replacement = "CA"))

ca_abbr


```

#### Example: Replacing Spiny with Delicious

```{r delicious}

delicious_inverts <- ca_abbr %>%  mutate(common_name = 
                                str_replace(common_name, 
                                            pattern = "spiny",
                                            replacement = "DELICIOUS"))


delicious_inverts
```
















